---
title: "jhu_ds2_w2_6"
author: "Sashin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## week 2 -- Dates and Time
### Dates and times in R
R has developed a special representation of dates and times:

* Date are represented by the *Date* class.
* Times are represented by the *POSIXct* or the *POSIXlt* class
* Dates are stored internally as the number of days since 1970-01-01
* Times are stored internally as the number of seconds since 1970-01-01

### Dates in R
Dates are represented by the Date class and can be coerced from a character string using the *as.Date()* function:
```{r}
x <- as.Date("1970-01-01")
x
unclass(x)
unclass(as.Date("1970-01-08"))
```
Dates before "1970-01-01" will be represented as negative numbers, since that date is 0. Dates are measured with respect to that date.

### Times in R
Times are represented using the *POSIXct* or the *POSIXlt* class, "POSIX" is a family of computing standards:

* *POSIXct* is just a very large integer under the hood; it is a useful class when you want to store times in something like a data frame.
* *POSIXlt* is a list underneath and it stores a bunch of other useful information like the day of the week, day of the year, month, day of the month.

There are a number of generic functions that work on dates and times:

* *weekdays*: give the day of the week
* *months*: give the month name
* *quarters*: give the quarter number ("Q1","Q2","Q3","Q4")

Times can be coerced from a character string using the *as.POSIXlt* or *as.POSIXct* function:
```{r}
X <- Sys.time()
x
p<- as.POSIXlt(X)
names(unclass(p))
p$sec
p$wday
```

You can also use the *POSIXct* format:
```{r}
x <- Sys.time()
x
unclass(x)
# x$sec  -- Error: $ operator is invalid for atomic vectors
p <- as.POSIXlt(x)
p$sec
```

Finally, there is the *strptime* function in case your dates are written in a different format.
```{r}
datestring <- c("January 10, 2012 10:40", "December 9, 2011 9:10")
x <- strptime(datestring, "%B %d, %Y %H:$M")
x
class(x)
```
Check docs for more details.


### Operations on Dates and Times
You can use mathematical operations on dates and times. Just + and -. You can also do comparisons (==, <=, etc).

```{r}
x <- as.Date("2012-01-01")
dates <- c("02/27/92", "02/27/92", "01/14/92", "02/28/92", "02/01/92")
times <- c("23:03:20", "22:29:56", "01:03:30", "18:21:03", "16:56:26")
y <- paste(dates, times)
strptime(y, "%m/%d/%y %H:%M:%S")
y
#x-y warning -- incompatible methods, non-numeric argument to binary operator

x<- as.POSIXlt(X)
z<- as.POSIXlt("2013-04-12")
x-z
```
It even keeps track of leap years, leap seconds, daylight savings, and time zones
use *tz* argument to specify timezones for *POSIXct* instances

Lots of plotting function can recognize datetime objects.


















