---
title: "functions2"
author: "Sashin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week 2
### Functions 2
#### Lazy Evaluation
Arguments to functions are evaluated lazily, so they are evaluated only as needed.
```{r}
f <- function(a,b){
  a*2
}
f(2)
```
This function never actually uses the argument b. So calling f(2) will not produce an error because the 2 gets positionally matched to a.



```{r}
h <- function(a,b){
  print(a)
  print(b)
}
#h(45)
```
*Error in print(b) : argument "b" is missing, with no default*

Notice that "45" got printed first before the error was triggered. This is because b did not have to be evaluated until after print(a). Once the function tried to evaluate print(b) it had to throw an error.




#### The ... Argument
This indicates a variable number of arguments that are usually passed on to other functions.

* ... is often used when extending another function and you don't want to copy the entire argument list of the original function

```{r}
myplot <-function(x,y, type="l", ...){
  plot(x,y,type=type, ...)
}
```
* Generic functions use ... so that extra arguments can be passed to methods (more later).

```{r}
mean
```
* The ... argument is necessary when the number of arguments passed to the function cannot be known in advance
```{r}
args(paste) # concatenates a set of strings, creates a vector of strings
args(cat)   # concatenates a string, prints to file
```
* One catch with ... is that any arguments that appear after it on the argument list must be named explicitly and cannot be partially matched.
```{r}
paste("a","b", sep=":")
paste("a","b", se=":")
```

