---
title: "jhu_w2_3"
author: "Sashin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week 2 -- Scoping Rules -- Symbol Binding
### Lexical scoping 
Why does all this matter?

* Typically, a function is defined in the global environment, so that the values of the free variables are found in the user's workspace.
* This behavior is logical for most people and is usually the *right thing* to do.
* However in R you can have functions defined *insided other functions*. Languages like C don't let you do this.
* Now things get interesting -- in this case the environment in which a function is defined in is the body of another function.


```{r}
make.power <- function(n){
  pow <- function(x){
    x*n
  }
}

cube <-make.power(3)
square <- make.power(2)
cube(3)
square(3)
```
These functions can be considered as constructor functions i.e. they create another function.


### Exploring a Function closure
Whats in the function's environment?
```{r}
ls(environment(cube))
get("n", environment(cube))
ls(environment(square))
get("n", environment(square))
```

Lexical vs Dynamic Scoping

```{r}
y <-10

f <- function(x){
  y <- 2
  y*2 + g(x)
}
g <- function(x){
  x*y
}
f(3)
```
* With lexical scoping, the value of y in the function g is looked up in the environment in which the function was defined, in this case the global environment, so the value of y is 10.
* With dynamic scoping, the value of y is looked up in the environment from which the function was called (sometimes referred to as the *calling environment*).
  + In R the calling environment is known as the *parent frame*.
* So the value of y would be 2

When a function is *defined* in the global environment and is subsequently *called* from the global environment, then the defining environment and the calling environment are the same. This can sometimes give the appearance of dynamic scoping.
```{r}
g <- function(x){ # x is a formal argument
  a <- 3  # local variable
  x+a+k   # k is a free variable
}
#g(2) Error in g(2) : object 'k' not found
k <-3
g(2)
```
other languages that support lexical scoping are Scheme, Perl, Python and Common Lisp (all languages that converge to Lisp)


### Consequences of lexical scoping

* In R, all objects must be stored in memory (hence large datasets can be a limiting factor)
* All functions must carry a pointer to their respective defining environments, which coulc be anywhere.
* In S-PLUS, free variables are always looked up in the global workspace, so everything can be stored on the disk because the *defining environmnet* of all functions is the same.
