---
title: "jhu_week2_2"
author: "Sashin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week 2 -- Scoping Rules -- Symbol Binding
### A Diversion on binding values to symbol

How does R know which value to assign to which symbol, When we type:
```{r}
lm <- function(x){x*x}
lm
```
how does R know what value to assign to the symbol lm? Why doesn't it give it the value of lm that is in the stats package?

When R tries to bind a value to a symbol, it searches through a series of *environments* to find the appropriate value. When you are working on the command line and need to retrieve the value of an R object, the order is roughly:

1. Search the global environment (current workspace i.e. everything you have created in the current session) for a symbol name matching the one requested.
2. Search the namespaces of each of the packages on the search list (all r packages currently loaded in to r)

The search list can be found via the *search* function:
```{r}
search()
```
There is an order to this search list.

### Binding values to Symbols

* The *global environment* or the user's workspace is always the first element of the search list and the *base* package is always the last.
* The order of packages on the search list matters!
* User's can configure which package gets loaded on start-up so you cannot assume that there will be a set list of packages available.
* When a user loads a package with *library* the namespace of that package gets put in position 2 of the search list (by default) and everything else gets shifted down the list.
* Note that R has separate namespaces for functions and non-functions so it's possible to have an object named *c* and a function named *c*


### Scoping Rules
The scoping rules for R are the feature that makes it different from the original S language.

* The scoping rules determine how a value is associated with a free variable in a function.
* R uses *lexical scoping* or *static scoping*. A common altenative is *dynamic scoping*.
* Related to the scoping rules is how R uses the search *list* to bind a value to a symbol.
* Lexical scoping turns out to particularly useful for simplifying statistical computations.

Consider the following function:
```{r}
i <- function(x,y){
  x^2 + y/z
}
```
This function has 2 formal arguments x and y. In the body of the function there is another symbol z. In this case z is called a *free variable*. The scoping rules of a language determine how values are assigned to free variables. Free variables are not formal arguments and are not local variables (assigned inside the function body).


### Lexical Scoping
Lexical scoping in R means that:
*the values of free variables are searched for in the environment in which the function was defined*

What is an environment?

* An *environment* is a collection of (symbol, value) pairs i.e. x is a symbol and 3.14 might be it's value.
* Every environment has a parent environment; it is possible for an environment to have multiple *children*
* the only environment without a parent is the empty environment
* A function + an environment = a *closure* or *function closure*

Searching for the value for a free variable:

* If the value of the symbol is not found in the environment in which a function was defined, then the search is continued in the *parent environment*
* The search continues down the sequence of parent environments until we hit the *top-level environment*, this is usually the global environment (workspace) or the namespace of a package.
* After the top-level environment, the search continues down the search list until we hit the *empty environment*. If a value for a given symbol cannot be found once the empty environment is arrived at, then an error is thrown.

