---
title: "jhu_r_w2_3"
author: "Sashin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week 2 -- Scoping Rules -- Lexical Scoping
### Application: Optimization
Why is any of this useful?

* Optimization routines in R like *optim*, *nim*, and *optimize* require you to pass a function whose argument is a vector of parameters (e.g. a log-likelihood)
* However, an object function might depend on a host of other things besides its parameters (like *data*)
* When writing software which does optimization, it may be desirable to allow the user to hold certain parameters  fixed.

### Maximizing a Normal Likelihood
Write a *constructor* function (creates an objective function)
```{r}
make.NegLokLik <-function(data, fixed=c(FALSE,FALSE)){
  params <- fixed
  function(p){
    parms[!fixed] <- p
    mu <- params[1]
    sigma <- params[2]
    a <- -0.5*length(data)*log(2*pi*sigma^2)
    b <- -sum((data-mu)^2) / (sigma^2)
    -(a+b)
  }
}
```
NOTE: Optimization functions in R *minimize* functions, so you need to use the negative log-likelihood

```{r}
set.seed(1); normals <- rnorm(100, 1, 2)
nLL <- make.NegLokLik(normals)
nLL
ls(environment(nLL))
```
When you define a function inside another one there will be a pointer to the parent function (environment).

"data" is set to be the value of normals which was passed to the constructor function.

### Estimating Parameters
```{r}
#optim(c(mu=0,sigma=1), nLL)

# Fixing ro to 2
#nLL <- make.NegLokLik(normals, c(FALSE,2))
#optimize(nLL, c(-1,3))$minimum
# Fixing mu to 1
#nLL <- make.NegLokLik(normals, c(1,FALSE))
#optimize(nLL, c(LE-6,10))$minimum
```


### Lexical scoping summary

* Objective functions can be *built* which contain all of the necessary data for evaluating the function.
* No need to carry around long argument lists - useful for interactive and exploratory work.
* Code can be simplified and cleaned up

* For more information, check out article *"Lexical Scope and Statistical Computing"* by Robert Gentlemen

